---
title: "P8106 Data Science Homework 1"
author: "Emil Hafeez"
geometry: "left=2.57cm,right=2.57cm,top=2.57cm,bottom=2.57cm"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
--- 

# Section 1: Setup

## Libraries and Options

```{r libraries, include = F}
library(lme4)
library(lmerTest)
library(tidyverse)
library(ggplot2)
library(plyr)
library(patchwork)
```

```{r opts chunk, include = F}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60))

knitr::opts_chunk$set(echo = TRUE)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```

```{r input}
train_dataframe = read_csv("./data/solubility_train.csv")

```


## Prompt

In this exercise, we will predict solubility of compounds using their chemical structures.

The training data are in the file “solubility train.csv” and the test data are in “solubil-
ity test.csv”. Among the 228 predictors, 208 are binary variables that indicate the presence

or absence of a particular chemical substructure, 16 are count features, such as the number
of bonds or the number of bromine atoms, and 4 are continuous features, such as molecular
weight or surface area. The response is in the column “Solubility” (the last column).

## Questions

(a) Fit a linear model using least squares on the training data and calculate the mean squared error using the test data.

(b) Fit a ridge regression model on the training data, with λ chosen by cross-validation. Report the test error.

(c) Fit a lasso model on the training data, with λ chosen by cross-validation. Report the test error and the number of non-zero coefficient estimates in your model.

(d) Fit a principle component regression model on the training data, with M chosen by cross-validation. Report the test error and the value of M selected by cross-validation.

(e) Which model will you choose for predicting solubility?

# Section 2: Implementation and Results
(a) Fit a linear model using least squares on the training data and calculate the mean squared error using the test data.

```{r}

```













